---
title: "A3: Jane Austen SPL Library Checkouts"
output: html_document
---

### Introduction

In this report, I will analyze the Seattle library checkouts of works by Jane Austen. Despite the time, Jane Austen continues to be a prevalent author today. Many cherish and admire Jane Austen's novels (me included). Her stories have been translated into several languages and have inspired multiple adaptations, documentaries, and retellings. Jane Austen has various works, including short stories and, of course, her novels. For this analysis, I decided to focus on her six novels: *Sense and Sensibility (1811)*, *Pride and Prejudice (1813*), *Mansfield Park (1814)*, *Persuasion (1817)*, and *Northanger Abbey (1817)*. I excluded *Sandition* because although it was published posthumously, like *Persuasion* and *Northanger Abbey*, it was uncompleted at the time of her death. 
 
I used data from the [SPL Open Data portal](https://data.seattle.gov/Community/Checkouts-by-Title/tmmm-ytt6/explore) and filtered out all the checkout entries for all materials with Jane Austen as the creator (this included entries that contained "Austen, Jane"). In addition, I included all the years possible to see the popularity trends over time for each novel. I did not exclude any material type but rather had all the different types, including ebooks, audiobooks, sound disks, and more. I wanted to understand how famous each story was as a whole; thus, I tried to capture the data from different forms. Furthermore, I was interested in discovering how the material type may have changed over the years.

### Summary Information
```{r, echo = FALSE, code = readLines("summary.R"), message = FALSE, warning = FALSE}
```
The data set revealed exciting details. For example, the data disclosed that in `r year_max_checkouts`, people checked out Jane Austen novels the most with `r max_checkouts` checkouts. This is a fascinating fact to process and shows the relevance of classics in today's age. I then took a step back to analyze the total checkouts from 2006 to 2023 and find which novel was the most popular. As expected, *`r popular_novel`* was checked out the most over the years. The novel (in different forms from physical to digital) was checked out a total of `r popular_novel_checkouts` times! At the same time, *`r unpopular_novel`* was the least popular, only checked out `r unpopular_novel_checkouts` times over the years. 

### The Dataset

The data was collected by the Seattle Public Library and published by the City of Seattle on its open data portal. The dataset captures monthly counts by the title of all checkouts from 2005 to the present (2023). The primary data set contains slightly over 42 million rows and 12 variables, which include: usage class (physical or digital), checkout type (vendor tool used to check out the item), material type (e.g., book, movie, music, etc.), check out year and month, number of checkouts, title, ISBN, creator/author, subjects, publisher, and publication year. 

According to the City of Seattle, the data comes from several current and historical sources. Information about digital item checkouts comes from the media vendors, which include Overdrive, hoopla, Freegal, and RBDigital. However, there are two primary sources for physical items: the Legrady artwork data archives (for works older than 2016) and Horizon ILS (2016 to present).

Libraries have a history of keeping records private to protect their members. However, the Seattle Public Library is part of Seattle's Open Data Program, a program dedicated to making data collected by the City of Seattle available to increase transparency, accountability, research, and more. Therefore, one ethical question to consider is whether the privacy of Seattle Public Library members is protected. 

Some possible limitation or issue with the data is how "messy" it is. For example, title names differed between different material types of the same title. The book *Pride and Prejudice* has inconsistent title names based on their material type and publisher. Some copies contained an introductionove or annotation, which was mentioned in the title. Another issue was the inclusion of novels not by Austen. Jane Austen has two stories that inspired parody novels (*Pride and Prejudice and Zombies* and *Sense and Sensibility and Sea Monsters*). The dataset included them because they listed Jane Austen as the creator; thus, I had to ensure to eliminate checkouts from these books. 

### Jane Austen Novels Checkout Rankings 

First, I wanted to derive a visual representation of *total* checkouts between 2005 and 2023 of all six Jane Austen novels to get a complete picture of the overall popularity of each title in the Seattle Public Library. Thus, I used a bar chart to compare how often each title was checked out. 

The chart emphasizes the domination of *Pride and Prejudice* over the other novels. But it also demonstrates just how considerable this novel's influence is. Additionally, it displays the ranking of the other novels. *Persuasion* is the second most popular, followed by* Emma*, *Sense and Sensibility*, *Mansfield Park*, and, lastly (as we calculated earlier), *Northanger Abbey*. The chart also demonstrates how the total checkouts of Jane Austen's *Persuasion*, *Emma*, and *Sense and Sensibility* are very close together. 

However, the last two novels in the chart are low especially compared to *Pride and Prejudice*. One thing that sets *Mansfield Park* apart from *Pride and Prejudice* is the protagonist Franny Price, who, unlike Elizabeth, is more timid. Meanwhile, *Northanger Abbey* is a satirical Gothic novel that differs from Austen's other books. These differences help us understand what characteristics Jane Austen and Seattle Public Library readers look for in a book and enjoy reading. These two novels also have fewer movie adaptions, which can explain the fewer checkouts. 

``````{r, echo = FALSE, code = readLines("chart1.R"), warning = FALSE, message = FALSE}
```

### Total Checkouts of Each Novel 

To observe the monthly checkouts trends of each novel over time, I grouped the dataset by the date and title of the book to find the monthly sum of checkouts and created a line chart. The graph shows interesting revelations. As we already know, *Pride and Prejudice* was the top novel checked out in the library for the majority of the time over the years. However, the line chart provides a better perception of the novel's popularity. We see a notable gap between *Pride and Prejudice* and the other titles. But we also see significant spikes in the past two years. The pandemic and books may be substantial contributors behind the increase in not only *Pride and Prejudice* but other titles as well. 
We can also observe that *Persuasion* surpassed *Pride and Prejudice* at one point and reaching its highest on July 2022. From prior knowledge, I connected this to the Netflix adaptation of *Persuasion* which highly influenced this influx of checkouts. When researching further, I discovered that around this time, TikTok also started its book club and started with *Persuasion*. We also see this happening to *Emma*, which was recently adapted into a movie. 

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart2.R"), message = FALSE}
```

### Pride and Prejudice Checkouts by Material Type 

Since *Pride and Prejudice* is Jane Austen's top novel, I wanted to dig deeper into what medium is checked out the most over the years. Therefore, I filtered the data set for only *Pride and Prejudice* checkouts and calculated the total checkouts of each material type and date. I again decided to use a line graph to display this data since it is the most appropriate visualization to reveal changes over time (see below). 

I was surprised at the checkout trends, which differed from my expectations. Initially, I expected to observe more book checkouts than the chart displayed. This expectation probably stemmed from my bias and preference for physical books rather than anything digital. However, upon reflecting on the data, I realized that it is logical for other mediums, such as audio books and ebooks, to have greater checkouts since digital items have multiple copies. In contrast, physical books are limited due to capacity measures. 

The line chart also demonstrates the advancement and acceptance of technology over time, as monthly physical book checkouts have decreased substantially from 2005 to the present day. One noteworthy observation is the high volume of audio books in this novel. We see a significant spike in 2019 and increases afterward. The popularity of audio books demonstrates the shift from physical books, to ebooks and finally audio books.

Seeing this trend, I wonder if audio books have become people's preference or if this is specifically just a trend in *Pride and Prejudice*. The shift to audio books may be due to how busy people are in today's age and also hustle culture. People are more inclined to multitask, and audio books allow them to do so.

``````{r, echo = FALSE, code = readLines("chart3.R"), warning = FALSE, message = FALSE}
```


